---
title: "a3_task3_satie_airame.Rmd"
author: "Satie Airame"
date: "11/1/2020"
output: html_document
---

## Santa Barbara Urchins

### Data: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(here)
library(janitor)
library(lubridate)
```

Read in the urchin.csv data and convert to lower_snake_case.

```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

#### Compare the distribution of sizes between purple and red urchins using faceted histogram and boxplot.

```{r}
ggplot(data = urchins, aes(x = size)) +
  geom_histogram(bins = 15) + 
  facet_wrap(~common_name)
```

#### Purple urchins demonstrate a slightly positively skewed size distribution (measured as diameter in cm). Red urchins demonstrate a normal distribution with slightly larger sizes than purple urchins.

```{r}
ggplot(data = urchins, aes(x = common_name, y = size)) +
  geom_boxplot()

```

#### The average diameter of purple urchins is smaller than red urchins. The standard deviation of diameter for red urchins is greater than for purple urchins.

#### Create a summary table of mean, median, standard deviation, and sample size of diameters of purple urchins and red urchins.

```{r}
urchin_stats <- urchins %>% 
  group_by(common_name) %>% 
  summarize(mean_size = mean(size), 
            median_size = median(size), 
            sd_size = sd(size), 
            n = n())
urchin_stats
```

#### The mean and median diameter of purple urchins is 4.5 cm (sd = 1.2). The mean size of red urchins is 6.8 cm and the median is 7.0 cm (sd = 2.0). The difference between mean urchin sizes is 2.3 cm.

### Filter purple_urchins, convert data to class "Date", create new column "year", create new column "site_full" with full name of sites.

```{r}
purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>%
  mutate(date_new = mdy(date)) %>% 
  mutate(year_new = year(date_new)) %>%
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"))

```

### Create summary table of total number purple urchins observed by site and year_new.

```{r}
purple_urchins_stats <- purple_urchins %>%
  count(site,year_new)
purple_urchins_stats
```

### Make exploratory graph of counts of purple urchins recorded at each site over time for the years in the study.

```{r}
ggplot(data = purple_urchins_stats, aes(x=year_new, y=n))+
  geom_line(aes(color = site))
```

#### The number of purple urchins at different sites over time. Urchins at all sites declined sinc 2008 and exhibited steep declines in 2011-2012 and 2019-2020.
